Management induced changes of soil organic carbon on global croplands

### 2020-07-14 SOCbudgetPaper/SOCbudget.tex
small update on written paper part

\title{Historical Soil Organic Carbon Budget}

\Author[1]{Kristine}{Karstens}
\Author[1]{Benjamin Leon}{Bodirsky}
\Author[1]{Alexander}{Popp}

\affil[1]{Potsdam-Institut of Climate Impacts Research, Potsdam,
Germany}

@@

\correspondence{Kristine\ Karstens\ (kristine.karstenst@pik-potsdam.de)}

@@

\begin{abstract}
SOC one of larges c sinks on earth (3 times larger biosphere pool).
Agricultural management leads to a depletion of soil organic carbon.
However this depletion of soil organic carbon (SOC) pools are so far not
well represented in global assessments of historic carbon emissions.
While SOC models often represent well the biochemical processes that
lead to the accumulation and decay of SOC, the management decisions
driving these biophysical processes are still little investigated. Here
we create a spatial explicit data set for crop residue and manure
management on cropland based on global historic production (FAOSTAT) and
land-use (LUH2) data and combine it with the IPCC Tier 2 approach to
create a half degree resolution soil organic carbon budget on mineral
soils. We estimate that due to arable farming soils have lost over (?)
GtOC of which (??) GtOC have been released within the period 1990-2010.
We show that, our results on global scale based on Tier 2 IPCC
methodolgy are in good agreement with Tier 1 default assumptions. We
also find that SOC is very sensitive to management decision such as
residue recycling indicating the nessessity to incorporated better
management data in soil models.
\end{abstract}

@@

\section{Method (50)}

We compiled calculations as open-source R packages available under
github.com/pik-piam/mrcommons (management related functions) and
github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which
are both based on github.com/pik-piam/madrat, a framework improving
reproducibility and transparency in data processing. In the following
chapter we outine the most important relationships and assumptions. See
table @ref(append:subsection2mrfunctions) for further information on
corresponding code within the R packages.

@@

\subsubsection{Steady-state SOC stocks and decay rates}

In a simple first order kinetic approach the steady-state soil organic
carbon stocks \(SOC^{eq}\) are given by \begin{equation}
SOC^{eq} =\frac{C^{\textrm{in}}}{k} \qquad\forall\quad i,t
\label{eq:inoutflow}
\end{equation} with \(C^{\textrm{in}}\) being carbon inputs to the soil,
\(k\) denoting the soil organic carbon decay rate; as well as \(i\)
representing grid cells indecies and \(t\) years. We use for our
calculations the steady-state method of the refinement of the IPCC
guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils, which
assume three soil carbon sub-pools (active, slow and passive) and
entangled dynamics between them. Annual carbon inflow to each sub-pool
(see @ref(sec:carboninputs)) and annual decay rates (see
@ref(sec:tier2)) of each sub-pool are still the key components to
determining steady-state SOC stocks.

\hypertarget{sec:carboninputs}{%
\paragraph{Carbon Inputs to the Soil}\label{sec:carboninputs}}

We account for different carbon input sources depending on our two
distinguished land-use types (see table @ref(tab:datasourceinputs)); in
particular for recycled crop residues, below ground biomass of crops
(for both see @ref(sec:residues)) and recycled manure (see
\ref(sec:livst\_manure)) on croplands; and litterfall
(\ref(\citet{LPJmL4_1})) on natural vegetation.

Following the IPCC methodology carbon inputs are disaggregated into
metabolic and structural components depending on their lignin and
nitrogen content (see @ref(ipcc\_2019\_2019)). For each component the
sum over all carbon input sources is allocated to the respective SOC
sub-pools via transfer coefficients. This implies that not only the
amount of carbon, but also their structural composition is determining
the effective inflow. Data sources for all considered carbon inputs as
well as for lignin and nitrogen content can be found in table
@ref(tab:datasourceinputs).

 \begin{table*}[h]
 \caption{Type and data sources for carbon inputs to different land-use types }
 \begin{tabular}{l l l l}
 \tophline
  \textbf{land-use types}   & \textbf{source of carbon inputs} & \textbf{data source} & \textbf{nitrogen and lignin content} \\
 \middlehline
 \multirow{3}{*}{Cropland} & residues & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2]  \\
                            & dead below ground biomass of crops & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2] \\
                            & manure & FAOSTAT, Isabelle [2, sec:manure] & default values given by [2] \\
                            \hline
  Natural vegetation        & annual litterfall & LPJmL4 [4]& \begin{minipage}[t]{0.28\columnwidth}\raggedright\strut Nitrogen and lignin content of tree compartments used in CENTURY [4] \strut \end{minipage}\tabularnewline
 \bottomhline
 \end{tabular}
 \label{tab:datasourceinputs}
 \belowtable{}
 \end{table*}

\hypertarget{sec:tier2}{%
\paragraph{Soil Organic Carbon decay (300)}\label{sec:tier2}}

The sub-pool specific decay rates are influenced by climatic conditions,
biophysical and biochemical soil properties as well as management
factors that all vary over time (t) and space (i). Following the
steady-state method of the refinement of the IPCC guidelines vol.~4
(\citet{ipcc_2019_2019}) for mineral soils we consider temperature
(temp), water (wat), sand fraction (sf) and tillage (till) effects to
account for spatial variation of decay rates. Thus \(k_{sub}\) is given
by

@@

For cropland we distinguish the effect of different tillage (see
@ref(\#sec:tillage)) and irrigation (see @ref(\#sec:irrigation))
practices on decay rates, whereas on natural vegetation, we assume
rainfed and non-tilled conditions. Data sources as well as considered
effects for each land-use types are shown in table
@ref(tab:datasourcedecay). To account for variations of decay rates
within each grid cell due to different tillage and irrigation regimes,
average rates based on area shares are calculated.

@@

                      & Soil disturbance & tillage & [sec:tillage] \\
 \bottomhline
 \end{tabular}
 \label{tab:datasourcedecay}
 \belowtable{}
 \end{table*}

\subsubsection{SOC transfer between land-use types}

@@

with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area
expansion for a given land-use type \(lut\). Note that \(!lut\) denotes
the sum over all other land-use types, which decreases in the specific
time step \(t\). Data sources and methodology on land-use states and
changes are described in @ref(sec:landuse).

\subsubsection{Total SOC stocks}

Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the
calculated steady-state stock \(SOC^{eq}\) for each land-use types
(lut), each sub-pool (sub) and each annual time step (t) as represented
in equation @ref(eq:steadystate).

\begin{equation}
SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall\quad\quad lut, sub, i.
\label{eq:steadystate}
\end{equation}

Reformulated equation \ref(eq:steadystate), we obtain a massbalance
equation as follows

\begin{equation}
SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation \ref(eq:inoutflow))}}  \qquad \forall\quad lut, sub, i.
\label{eq:steadystate}
\end{equation}

@@

\subsubsection{Initialisation of SOC pools}

@@

\paragraph{Crop Production}

Using half-degree yield data from LPJmL (\citep{LPJmL4_1}) as well as
half-degree cropland patterns (see @ref(\#sec:landuse)) we compile crop
group specific half-degree production patterns. We calibrate cellular
yields with one country-level calibration factor for each crop group to
meet historical FAOSTAT production (\citep{FAOSTAT}). Note that by using
physical cropland areas we account for multiple crop harvest events as
well as for fallows.

\paragraph{Crop Residue Production}

Crop residue production and management is based on a revised methodology
of (\citep{bodirsky2012}) and will be explained in key aspects again due
to its central role for soil carbon modelling. Starting from crop
production (\(CP\)) estimates of the harvested organs and their
respective harvested crop area (\(CA\)), we estimate above-ground
(\(AGR\)) and below-ground (\(BGR\)) residual biomass using crop group
(\(cg\)) specific harvest indices (\(HI\)) and root:shoot ratios
(\(RS\)) as follows

@@

\subparagraph{Burned Residues}

A fixed share of the AGR is assumed to be burned on field depending on
the per-capita income of the country. Following \citep{smil1999}) we
assume 25\% burn share for low-income countries according to worldbank
definitions (\(<\,1000\,\tfrac{USD}{yr}\)), 15\% for high-income
(\(>\,10000\,\tfrac{USD}{yr}\) and linearly interpolate shares for all
middle-income countries depending on their per-capita income. Depending
on the crop group 80--90\% of the residue carbon burned on the fields
are lost within the combustion process (\citep{ipcc_2006_2006}).

\subparagraph{Residue Usage}

We compile out of our 17 crop groups, three used residue groups (straw,
high-lignin and low-lignin residues) with additional demand for other
purposes and one residues with no double use (see
@ref(append:Table\_kcr2kres)). Residue feed demand for five different
livestock groups is based on country- and residue-group-specific feed
basekts (see \citep{weindl}) taking available AGR biomass as well as
livestock productivity into account. We estimate a material-use share
for the straw residues group of 5\% and a fuel-share of 10\% for all
used residues groups in low income countries according to worldbank
definitions (\(<\,1000\,\tfrac{USD}{yr}\)). For high-income
(\(>\,10000\,\tfrac{USD}{yr}\) no withdrawl for material or fuel use is
assumend, leaving middle-income countries with linearly interpolate
shares depending on their per-capita income. The remaining AGR as well
as all BGR are assumend to be recycled to the soil. We cut high
recycling shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to corrected for
outliers.

\paragraph{Dry Matter to Carbon Transformation}

To transform dry matter estimates into carbon, we compiled crop-group
and plant part specific carbon to dry matter (c:dm) ratios (see
@ref(append:Table\_c2dm)) (\citet{Ma2018}).

@@

\paragraph{Livestock Distribution}

To disaggregate country level FAOSTAT livestock production values to
half-degree pattern, we use the following rule based assumptions which
were inspired by the approach of \citep{gilbert} and uses feed basket
assumptions based on a revised methodology of \citet{weindl.We}
differentiate ruminant and monogastric systems, as well as extensive and
an intensive systems. For ruminants, we assume that livestock is located
at the production spot sof the feed intake. We distingush between grazed
pasture, which is converted into livestock products from extensive
systems; and all other (crop-based) feed stuff, which we consider to be
consumend in intensive systems. For poultry, egg and monogastric meat
production we use the per-capita income of the country to divide into
intensive and extensive production systems. For low-income countries
according to worldbank definitions (\textless1000 USD/yr), we assume
extensive production systems. We locate them according to built-up areas
shares based on the idea that these animals are held in households,
subsistence or small-holder farming systems with a high labour per
animal ratio. Intensive production is distributed within a country using
again crop-based feed production share, assuming that feed availability
is the most driving factor for livestock location.

\paragraph{Manure Excretion, Storage and Recycling}

Manure production and management is based on a revised methodology of
(\citep{bodirsky2012}) and will be explained in key aspects again due to
its central role for soil carbon modelling. Based on the gridded
livestock distribution we calculate excretions by estimating the
nitrogen balance of the livestock system on the basis of comprehensive
livestock feed baskets (\citep{weindl}), assuming that all nitrogen in
protein feed intake, minus the nitrogen in the slaughter mass, is
excreted. Carbon in excreted manure is estimated by applying fixed C:N
ratios (given by \citep[(][]{ipcc_2019_2019}). Depending on the feed
system we assume manure to be handled in four different ways: All manure
orginated from pasture feed intake is excreted directly to pastures and
rangelands (pasture grazing), deducting manure collected as fuel. Manure
fuel shares are estimated using IPCC default values
(\citet{ippc_2006_2006}). Whereas for low-income countries according to
worldbank definitions (\textless1000 USD/yr), we adopt a share of 25\%
of crop residues in feed intake directly consumend and excreted on crop
fields (stubble grazing), we do not consider any stubble grazing in
high-income countries (\(>\,10000\,\tfrac{USD}{yr}\), leaving
middle-income countries with linearly interpolate shares depending on
their per-capita income. For all other feed items we assume the manure
to be stored in animal waste management systems associated to animal
houses. To estimate the carbon actually recycled to the soil, we account
for carbon losses during storage and recycling shares in different
animal waste management and grazing systems. Whereas we assume no losses
for pasture and stubble grazing, we consider that the manure collected
as fuel is not recycled. For manure stored in different animal waste
management system we compiled carbon loss rates partly depending on the
nitrogen loss rates as specified in \citep{bodirsky2012} (see
@ref(append:Table\_clossAWMS))). We cut high application shares at
\(10\tfrac{\unit{tC}}{\unit{ha}}\) to corrected for outliers.

\subsubsection{Irrigation (100)}

@@

Additionally to the tier 2 approach of the refinement of IPCC guidelines
vol.~4 (\citet{ipcc_2019_2019}) and the detailed analysis of management
data coming with it, SOC changes can be estimated using the IPCC tier 1
approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006},
\citet{ipcc_2019_2019}). Here, stocks are calculated via stock change
factors (SCF) given by the IPCC for the topsoil (0-30 cm) and based on
observational data. SCF are differentiated by different crop and
management systems (\(m\)) reflecting different dynamics under changed
in- and outflows without explicitly tracking these. Moreover SCF vary
for different climatic zones (\(c\)) specified by the IPCC (see
@ref(append:climatemap)). The actual SOC stocks as thus calulated based
on a given reference SOC stock by

@@

@@

\section{Results}

@@

\includegraphics[width=16cm]{../ResultNotebooks/Output/Images/4panelfigure}
\newpage

\section{Discussion}

@@

\begin{itemize}
\item
  SOC initialization is bad (Task: find a better representation)
\item
  natural land representation is lacking proper parametrization of
  natural input (n, lg content of litterfall) (--\textgreater{} ask
  Christop again)
\item
  boreal zone and dry regions are not well represented (maybe because of
  the bad parameterization of the soil model for natural soils)
\item
  As pointed out by Keel (2017), Smith (2019) the results might be
  highly sensitive to carbon input calculations more precisly to below
  and above ground residue carbon estimates derived from harvested
  quantities. It has been questioned, that below ground residues might
  increase with a fixed root:shot ratio (maybe specifically in high end
  farming systems (?)). Following this argumentation SOC results shown
  in this study might especially in high-yielding farming system (europe
  etc.) overestimate actual SOC stocks.
\item
  fertilizer interaction (--\textgreater{} ask LPJmLer again)
\item
  diaggregation of manure with urban area is leeding to a lot of
  displaced manure (?) that is cut off
\item
  mismatches between FAO/LUH
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontribution{Karstens wrote code and paper build on work of
Bodirsky. Bodirsky and Popp revised paper.} %% optional section

>>>>

\title{Management induced changes of soil organic carbon on global croplands \newline or \newline How management changed soil organic carbon on global croplands \newline or \newline Agricultural soil have lost 13 GtC topsoil carbon \newline or \newline Linking agricultural management data to soil modeleling -- a global approach}

\Author[1]{Kristine}{Karstens}
\Author[1]{Benjamin Leon}{Bodirsky}
\Author[2]{Your}{Name}
\Author[1]{Alexander}{Popp}

\affil[1]{Potsdam-Institut of Climate Impacts Research, Potsdam, Germany}
\affil[2]{Your affiliation}

@@

\correspondence{Kristine\ Karstens\ (\href{mailto:kristine.karstens@pik-potsdam.de}{\nolinkurl{kristine.karstens@pik-potsdam.de}})}

@@

\begin{abstract}
Soil organic carbon (SOC) is one of larges c stocks on earth (3 times larger biosphere pool).
\end{abstract}

@@

\hypertarget{method-50}{%
\section{Method (50)}\label{method-50}}

We compiled calculations as open-source R packages available under github.com/pik-piam/mrcommons (management related functions) and github.com/pik-piam/mrSOCbudget (soil dynamic related functions), which are both based on github.com/pik-piam/madrat, a framework improving reproducibility and transparency in data processing.
In the following chapter we outine the most important relationships and assumptions. See table \ref{append:subsection2mrfunctions} for further information on corresponding code within the R packages.

@@

\hypertarget{steady-state-soc-stocks-and-decay-rates}{%
\subsubsection{Steady-state SOC stocks and decay rates}\label{steady-state-soc-stocks-and-decay-rates}}

In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{eq}\) are given by
\begin{equation}
SOC^{eq} =\frac{C^{\textrm{in}}}{k} \qquad\forall\quad i,t
\label{eq:inoutflow}
\end{equation}
with \(C^{\textrm{in}}\) being carbon inputs to the soil, \(k\) denoting the soil organic carbon decay rate; as well as \(i\) representing grid cells indecies and \(t\) years. We use for our calculations the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils, which assume three soil carbon sub-pools (active, slow and passive) and entangled dynamics between them. Annual carbon inflow to each sub-pool (see \ref{sec:carboninputs}) and annual decay rates (see \ref{sec:tier2}) of each sub-pool are still the key components to determining steady-state SOC stocks.

\hypertarget{carbon-inputs-to-the-soil-seccarboninputs}{%
\paragraph*{Carbon Inputs to the Soil \{\#sec:carboninputs\}}\label{carbon-inputs-to-the-soil-seccarboninputs}}
\addcontentsline{toc}{paragraph}{Carbon Inputs to the Soil \{\#sec:carboninputs\}}

We account for different carbon input sources depending on our two distinguished land-use types (see table \ref{tab:datasourceinputs}); in particular for recycled crop residues, below ground biomass of crops (for both see \ref{sec:residues}) and recycled manure (see @ref(sec:livst\_manure)) on croplands; and litterfall (\citep{LPJmL4_1}) on natural vegetation.

Following the IPCC methodology carbon inputs are disaggregated into metabolic and structural components depending on their lignin and nitrogen content (see \citep{ipcc_2019_2019}). For each component the sum over all carbon input sources is allocated to the respective SOC sub-pools via transfer coefficients. This implies that not only the amount of carbon, but also their structural composition is determining the effective inflow. Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \ref{tab:datasourceinputs}.

 \begin{table*}[h]
 \caption{Type and data sources for carbon inputs to different land-use types }
 \begin{tabular}{l l l l}
 \tophline
  \textbf{land-use types}   & \textbf{source of carbon inputs} & \textbf{data source} & \textbf{nitrogen and lignin content} \\
 \middlehline
 \multirow{3}{*}{Cropland} & residues & FAOSTAT, LPJmL4 [2, \\ref{sec:residues}] & default values given by [2]  \\
                            & dead below ground biomass of crops & FAOSTAT, LPJmL4 [2, \\ref{sec:residues}] & default values given by [2] \\
                            & manure & FAOSTAT, Isabelle [2, sec:manure] & default values given by [2] \\
                            \hline
  Natural vegetation        & annual litterfall & LPJmL4 [4]& \begin{minipage}[t]{0.28\columnwidth}\raggedright\strut Nitrogen and lignin content of tree compartments used in CENTURY [4] \strut \end{minipage}\tabularnewline
 \bottomhline
 \end{tabular}
 \belowtable{}
 \label{tab:datasourceinputs}
 \end{table*}

\hypertarget{soil-organic-carbon-decay-300-sectier2}{%
\paragraph*{Soil Organic Carbon decay (300) \{\#sec:tier2\}}\label{soil-organic-carbon-decay-300-sectier2}}
\addcontentsline{toc}{paragraph}{Soil Organic Carbon decay (300) \{\#sec:tier2\}}

The sub-pool specific decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that all vary over time (t) and space (i). Following the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citep{ipcc_2019_2019}) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to account for spatial variation of decay rates. Thus \(k_{sub}\) is given by

@@

For cropland we distinguish the effect of different tillage (see \ref{sec:tillage}) and irrigation (see \ref{sec:irrigation}) practices on decay rates, whereas on natural vegetation, we assume rainfed and non-tilled conditions. Data sources as well as considered effects for each land-use types are shown in table \ref{tab:datasourcedecay}. To account for variations of decay rates within each grid cell due to different tillage and irrigation regimes, average rates based on area shares are calculated.

@@

                      & Soil disturbance & tillage & [sec:tillage] \\
 \bottomhline
 \end{tabular}
 \belowtable{}
 \label{tab:datasourcedecay}
 \end{table*}

\hypertarget{soc-transfer-between-land-use-types}{%
\subsubsection{SOC transfer between land-use types}\label{soc-transfer-between-land-use-types}}

@@

with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area expansion for a given land-use type \(lut\). Note that \(!lut\) denotes the sum over all other land-use types, which decreases in the specific time step \(t\). Data sources and methodology on land-use states and changes are described in \ref{sec:landuse}.

\hypertarget{total-soc-stocks}{%
\subsubsection{Total SOC stocks}\label{total-soc-stocks}}

Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the calculated steady-state stock \(SOC^{eq}\) for each land-use types (lut), each sub-pool (sub) and each annual time step (t) as represented in equation \eqref{eq:steadystate}.

\begin{equation}
SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall\quad\quad lut, sub, i.
\label{eq:SOCstate}
\end{equation}

Reformulated equation \eqref{eq:SOCstate}, we obtain a massbalance equation as follows

\begin{equation}
SOC_{t} = SOC_{t-1} - \underbrace{SOC_{t-1} \cdot k_{t} \cdot 1\unit{a}}_{\text{outflow}} + \overbrace{SOC^{eq}_{t} \cdot k_{t} \cdot 1\unit{a}}^{\text{input (using equation \\eqref{eq:inoutflow})}}  \qquad \forall\quad lut, sub, i.
\label{eq:steadystate2budget}
\end{equation}

@@

\hypertarget{initialisation-of-soc-pools}{%
\subsubsection{Initialisation of SOC pools}\label{initialisation-of-soc-pools}}

@@

\textbf{Crop Production}
Using half-degree yield data from LPJmL (\citep{LPJmL4_1}) as well as half-degree cropland patterns (see \ref{sec:landuse}) we compile crop group specific half-degree production patterns. We calibrate cellular yields with one country-level calibration factor for each crop group to meet historical FAOSTAT production (\citep{FAOSTAT}). Note that by using physical cropland areas we account for multiple crop harvest events as well as for fallows.

\textbf{Crop Residue Production}
Crop residue production and management is based on a revised methodology of (\citep{bodirsky2012}) and will be explained in key aspects again due to its central role for soil carbon modelling. Starting from crop production (\(CP\)) estimates of the harvested organs and their respective harvested crop area (\(CA\)), we estimate above-ground (\(AGR\)) and below-ground (\(BGR\)) residual biomass using crop group (\(cg\)) specific harvest indices (\(HI\)) and root:shoot ratios (\(RS\)) as follows

@@

\textbf{Burned Residues}
A fixed share of the AGR is assumed to be burned on field depending on the per-capita income of the country. Following \citep{smil1999}) we assume 25\% burn share for low-income countries according to worldbank definitions (\(<\,1000\,\tfrac{USD}{yr}\)), 15\% for high-income (\(>\,10000\,\tfrac{USD}{yr}\) and linearly interpolate shares for all middle-income countries depending on their per-capita income. Depending on the crop group 80--90\% of the residue carbon burned on the fields are lost within the combustion process (\citep{ipcc_2006_2006}).

\textbf{Residue Usage}
We compile out of our 17 crop groups, three used residue groups (straw, high-lignin and low-lignin residues) with additional demand for other purposes and one residues with no double use (see @ref(append:Table\_kcr2kres)). Residue feed demand for five different livestock groups is based on country- and residue-group-specific feed basekts (see \citep{weindl}) taking available AGR biomass as well as livestock productivity into account. We estimate a material-use share for the straw residues group of 5\% and a fuel-share of 10\% for all used residues groups in low income countries according to worldbank definitions (\(<\,1000\,\tfrac{USD}{yr}\)). For high-income (\(>\,10000\,\tfrac{USD}{yr}\) no withdrawl for material or fuel use is assumend, leaving middle-income countries with linearly interpolate shares depending on their per-capita income. The remaining AGR as well as all BGR are assumend to be recycled to the soil. We cut high recycling shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to corrected for outliers.

\textbf{Dry Matter to Carbon Transformation}
To transform dry matter estimates into carbon, we compiled crop-group and plant part specific carbon to dry matter (c:dm) ratios (see @ref(append:Table\_c2dm)) (\citet{Ma2018}).

@@

\textbf{Livestock Distribution}
To disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions which were inspired by the approach of \citep{gilbert} and uses feed basket assumptions based on a revised methodology of \citet{weindl.We} differentiate ruminant and monogastric systems, as well as extensive and an intensive systems.
For ruminants, we assume that livestock is located at the production spot sof the feed intake. We distingush between grazed pasture, which is converted into livestock products from extensive systems; and all other (crop-based) feed stuff, which we consider to be consumend in intensive systems.
For poultry, egg and monogastric meat production we use the per-capita income of the country to divide into intensive and extensive production systems. For low-income countries according to worldbank definitions (\textless1000 USD/yr), we assume extensive production systems. We locate them according to built-up areas shares based on the idea that these animals are held in households, subsistence or small-holder farming systems with a high labour per animal ratio. Intensive production is distributed within a country using again crop-based feed production share, assuming that feed availability is the most driving factor for livestock location.

\textbf{Manure Excretion, Storage and Recycling}
Manure production and management is based on a revised methodology of (\citep{bodirsky2012}) and will be explained in key aspects again due to its central role for soil carbon modelling. Based on the gridded livestock distribution we calculate excretions by estimating the nitrogen balance of the livestock system on the basis of comprehensive livestock feed baskets (\citep{weindl}), assuming that all nitrogen in protein feed intake, minus the nitrogen in the slaughter mass, is excreted. Carbon in excreted manure is estimated by applying fixed C:N ratios (given by \citep[(][]{ipcc_2019_2019}).
Depending on the feed system we assume manure to be handled in four different ways:
All manure orginated from pasture feed intake is excreted directly to pastures and rangelands (pasture grazing), deducting manure collected as fuel. Manure fuel shares are estimated using IPCC default values (\citet{ippc_2006_2006}).
Whereas for low-income countries according to worldbank definitions (\textless1000 USD/yr), we adopt a share of 25\% of crop residues in feed intake directly consumend and excreted on crop fields (stubble grazing), we do not consider any stubble grazing in high-income countries (\(>\,10000\,\tfrac{USD}{yr}\), leaving middle-income countries with linearly interpolate shares depending on their per-capita income.
For all other feed items we assume the manure to be stored in animal waste management systems associated to animal houses.
To estimate the carbon actually recycled to the soil, we account for carbon losses during storage and recycling shares in different animal waste management and grazing systems. Whereas we assume no losses for pasture and stubble grazing, we consider that the manure collected as fuel is not recycled. For manure stored in different animal waste management system we compiled carbon loss rates partly depending on the nitrogen loss rates as specified in \citep{bodirsky2012} (see @ref(append:Table\_clossAWMS))). We cut high application shares at \(10\tfrac{\unit{tC}}{\unit{ha}}\) to corrected for outliers.

\hypertarget{irrigation-100}{%
\subsubsection{Irrigation (100)}\label{irrigation-100}}

@@

Additionally to the tier 2 approach of the refinement of IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) and the detailed analysis of management data coming with it, SOC changes can be estimated using the IPCC tier 1 approach of IPCC guidelines vol.~4 (\citet{ipcc_2006_2006}, \citet{ipcc_2019_2019}). Here, stocks are calculated via stock change factors (SCF) given by the IPCC for the topsoil (0-30 cm) and based on observational data. SCF are differentiated by different crop and management systems (\(m\)) reflecting different dynamics under changed in- and outflows without explicitly tracking these. Moreover SCF vary for different climatic zones (\(c\)) specified by the IPCC (see \ref{append:climatemap}). The actual SOC stocks as thus calulated based on a given reference SOC stock by

@@

According to \citep{ipcc_2019_2019} the formulation of absolute changes like
might be neccessary as well.

@@

\hypertarget{results}{%
\section{Results}\label{results}}

We present simulation results of the SOC budget of the period from 1975 to 2010 mainly focusing on cropland areas. Detailed total global and regional results of the state of the worlds SOC in comparison to other estimates can be found in the supplement.

\hypertarget{soc-distribution-and-depletion}{%
\subsection{SOC distribution and depletion}\label{soc-distribution-and-depletion}}

Figure

\begin{figure}
\includegraphics[width=18cm]{../ResultNotebooks/Output/Images/4panelfigure} \caption{(a) ... , (b) ... , (c) ... , (d) ... .}\label{fig:SOCmaps}
\end{figure}

\hypertarget{comparison-of-stock-changes-factors}{%
\subsection{comparison of stock changes factors}\label{comparison-of-stock-changes-factors}}

@@

\hypertarget{global-agricultrual-carbon-flows-and-stocks}{%
\subsection{global agricultrual carbon flows and stocks}\label{global-agricultrual-carbon-flows-and-stocks}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

\begin{itemize}
\item
  SOC initialization is bad (Task: find a better representation)
\item
  natural land representation is lacking proper parametrization of natural input (n, lg content of litterfall) (--\textgreater{} ask Christop again)
\item
  boreal zone and dry regions are not well represented (maybe because of the bad parameterization of the soil model for natural soils)
\item
  As pointed out by Keel (2017), Smith (2019) the results might be highly sensitive to carbon input calculations more precisly to below and above ground residue carbon estimates derived from harvested quantities. It has been questioned, that below ground residues might increase with a fixed root:shot ratio (maybe specifically in high end farming systems (?)). Following this argumentation SOC results shown in this study might especially in high-yielding farming system (europe etc.) overestimate actual SOC stocks.
\item
  fertilizer interaction (--\textgreater{} ask LPJmLer again)
\item
  ``The Tier 2 method does not simulate C change but simply calculates an annual C stock change from the
  current C stock to the future steady-state soil C stock calculated based on current conditions.'' Leading to the fact that our total stock results are highly uncertained.
\item
  diaggregation of manure with urban area is leeding to a lot of displaced manure (?) that is cut off
\item
  mismatches between FAO/LUH
\end{itemize}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontribution{Karstens wrote code and paper build on work of Bodirsky. Bodirsky, and Popp revised paper.} %% optional section

### 2020-07-14 SOCbudgetPaper/discussion.Rmd
update method part + smaller stuff in other section and result figure
small update on written paper part

@@

- pasture dynamics are neglected and treated as natveg

>>>>

- "The Tier 2 method does not simulate C change but simply calculates an annual C stock change from the
current C stock to the future steady-state soil C stock calculated based on current conditions." Leading to the fact that our total stock results are highly uncertained.

@@

- pasture dynamics are neglected and treated as natveg  / Note that manure excreted to pastures is neglected within these analysis, since we focus purely on cropland dynamics.
